Clouds and Sensitivities Across a Hierarchy of GFDL CMIP6 Models

### 2020-08-25 Silvers_cfmip.tex
added the reference for Soden et al., 2004 and minor edits to the CRE discussion.

			% Activate to display a given date or no date
%\title{Using the Walker Circulation to connect low-level cloud changes to deep convective entrainment}
\title{Clouds and Sensitivities Across a Hierarchy of GFDL CMIP6 Models}
\authors{Levi G. Silvers, David Paynter, V. Balaji, Chris Blanton, Chris Dupuis, Jeffrey Durachta, Huan Guo, 
Isaac Held, Richard Hemmler, Jasmin G. John, Pu Lin, Colleen McHugh, Serguei Nikonov, Jeffrey Ploshay, Aparna Radhakrishnan, 
Kristopher Rand, Thomas Robinson, Hans Vahlenkamp, Chandin Wilson, Bruce Wyman, Yujin Zeng, Ming Zhao}
% Isaac Held?  Michael Winton?  Pu Lin? Songmiao Fan?  M. Daniel Schwarzkopf?

@@

\begin{figure}
  \centering
% figure generated with calip25_isccp_am4_totcld.ncl, figurenumber = 1
  \includegraphics[width=0.6\columnwidth]{cfmipfigs/ISCCP_CALIPSO_clt_correctedBiasPattern_nogrid.eps}
  %\includegraphics[width=0.75\columnwidth]{cfmipfigs/clt_isccp_calipso_am4_4pan.eps}
  \caption{Comparison of total cloud fraction from the AM4 \textit{amip} experiment with two observational 
  data sets: ISCCP (left) and CALIPSO (right).  The top row shows the observational data.  The bottom row
  shows the result of subtracting the observations from the AM4.0 data.  Numbers in the title above each 
  panel give the global mean cloud fraction (top), and the bias of global mean cloud fraction (bottom)}
  \label{fig:clt_isccp_calipso}
\end{figure}

@@

Before discussing the climate sensitivities of AM4/CM4 and how they are influenced by cloud feedbacks we assess
the clouds and radiative fluxes which are simulated by AM4.0 over the recent period of 1979-2014.  This is done 
by comparing satellite based observations to the \textit{amip} experiment that is part of GFDLs 
contribution to CMIP6.  
During the development and documentation of AM4.0 \cite{Zhao_etal_2018a, Zhao_etal_2018b}, 
(Z18a and Z18b hereafter) 
and CM4 \cite{Held_etal_2019}, specific cloud properties such as the 
cloud droplet number, cloud fraction, and the partition of cloud water into liquid or ice were noted, but 
accurately modeling the radiative fluxes at the top of the atmosphere was determined to be a 
higher priority partly because of the large observational uncertainties of those cloud properties (Z18a,b).   
Based on the initial analysis for those papers, the cloud properties of AM4.0 were determined to be
at least as consistent with observations as previous generations of GFDL models.  Z18a also noted 
that AM4.0 underestimates the liquid droplet number over oceans and downwind of the aerosol emission 
sources.   
Comparisons of AM4.0 with observations from ISCCP and MODIS showed an underestimate of 
the optically thin low clouds and the mid-level clouds (Z18a, Figure 10).  
However, the observational uncertainty for some of these properties prohibited firm conclusions.  This paper
uses additional observations and a more thorough analysis of the clouds in AM4/CM4 to determine 
the quality of the cloud simulations.        
In this section we first compare the simulated and observed cloud fraction.  We then compare  
the simulated TOA radiative fluxes from AM4.0 to those observed by CERES.    
Throughout the paper we utilize what has come to be known as the Cloud Radiative Effect (CRE).  The CRE is the 
difference between the full-sky and clear-sky radiative fluxes (this quantity is also sometimes referred
to in the literature as the cloud radiative forcing, CRF).

\begin{figure}
  \centering
% figure generated with calip25_isccp_am4_totcld.ncl, figurenumber = 3
  \includegraphics[width=0.6\columnwidth]{/Users/silvers/Research/Silvers_cfmip_paper/cf_misr_cldfrac_below3km.eps}
  \caption{Observed and modeled clouds below 3km.  Clouds shown are divided into thick 
  (optical depth greater than and equal to 3.6) and thin (optical depth less than 3.6) 
  clouds.  The observations (bottom) are from the MISR satellite (2000-2019) and the modeled clouds (top) were derived 
  with the MISR simulator (amip experiment). Mean cloud fraction values for MISR 
  (AM4) are 18\% (13\%) for thick clouds and 22\% (6\%) for thin clouds}
  \label{fig:cf_misr_lowcl}
\end{figure}
% MISR observations are from March 2000 through November 2019 and came from Roger Marchand

\subsection{COSP}
In order to compare the observed and modeled clouds as consistently as possible across multiple platforms 
GFDL has implemented a collection of satellite simulators which are part of the 
CFMIP (Cloud Feedback Model Intercomparison Project; see \cite{Webb_etal_2017}).  Satellite simulator
are software tools that are used in coordination with GCMs to allow for model variables 
to be translated to variables that are more directly comparable with what satellites observe.  
These simulators are often referred to as the CFMIP Observation Simulator Package (COSP); for
details see Bodas-Salcedo et al. 2011.  This paper highlights simulators of the algorithms for 
four satellite datasets: ISCCP (the International Satellite Cloud Climatology Project), 
MODIS (the Moderate Resolution Imaging Spectroradiometer), 
MISR (the Multiangle Imaging Spectro-Radiometer), and 
CALIPSO (the Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation; 
Chepfer et al., 2010; Cesana and Chepfer, 2013).  By using multiple complimentary satellite 
datasets and simulators we can benefit from the different strengths of each satellite while largely 
minimizing the known problems and discrepancies of the observational data sets.

\begin{figure}
  \centering
% figure generated with cloud_table_AM4fig.ncl
  \includegraphics[width=0.98\columnwidth]{cfmipfigs/Silvers_cfmip_MISRv7_cloudtable.pdf}
  \caption{Cloud Fraction (colors and numbers (\%)) in joint histograms showing the variations of cloud fraction 
  as a function of cloud top height and optical depth (tau).
  The top panels show observations from MISR, bottom panels show MISR simulator output from the 
  \textit{amip} experiment.  Shown from left to right are the mean values for the full globe, between 30S-60S,
  between 30S-30N, and between 30N-60N.}
  \label{fig:misr_sim_vs_mod}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[angle=90,width=0.85\columnwidth]{cfmipfigs/calip_9pan_amip_obs_vs_AM4.eps}
  \caption{Cloud Fraction observed by CALIPSO (2007-2016) and modeled with the CALIPSO satelite simulator 
  (amip experiment).
  Dashed line shows meridional profiles of AM4.0 cloud fraction and solid line shows
  the meridional profile of observed cloud fraction.  With only a few exceptions in the polar regions, AM4 underestimates
  clouds in all three vertical ranges with the largest underestimate being of low-level clouds between +/-30.}
  \label{fig:calipso_9pan}
\end{figure}

@@

\section{The Role of Clouds in AM4.0/CM4.0/Aquaplanets in Determining and Influencing Climate Sensitivity}

@@

The set of new models developed at NOAA's GFDL that have participated in CMIP6 have a large range of climate sensitivities.  
One way of quantifying this sensitivity is with the climate feedback parameter, which 
varies among a suite of CMIP6 experiments between -0.46 and -2.1 $\rm{W m^{-2}K^{-1}}$ (Table 1).  
The models that are the focus of this paper (AM4.0 Aquaplanet, AM4.0, CM4.0) simulate several 
distinct configurations which account for some of the differences in the climate feedback parameter.  
After a brief review of the methods for estimating the climate sensitivity we show in this section how much of the difference is 
due to experimental configuration and how much is due to differences in the CRE.  
We also show how the cloud fraction varies among
the experiments as a function of latitude and consider the cloud distribution in the context of the observed
CRE from the previous section.  The CRE is an effective tool for evaluating the influence of 
clouds on the climate.  It is well known (Soden et al., 2004) that a change in the CRE is not a direct measure of the cloud 
feedback, and that non-cloud related values of CRE can occur.  However, there are many cases of 
the productive use of the CRE as a way to diagnose cloud effects (e.g. Cess et al., 1989; Webb et al., 2001; 
\citeA{Ringer_etal_2014}; Andrews et al., 2015), it is simple to calculate, and
it allows for direct comparisons with previous studies.  The true cloud feedback usually corresponds 
to a positive shift of the change in CRE values by roughly 0.3 $\rm{W m^{-2}K^{-1}}$ (Soden et al., 2004).  The total cloud feedback of a model 
is therefore more positive than will be implied by the CRE.

Our hypothesis for this study is that the large range of climate sensitivities that are found within this 
group of models is in large part due to variations in the clouds (both their climatology and their feedbacks).
The clouds in these simulations respond to a variety of different conditions including different
patterns of prescribed SST, the presence or absence of land/sea ice, and in the case of the aquaplanet
experiments, a lack of cloud-aerosol interactions which are present in the AGCM and coupled 
experiments (AOGCM).

\subsection{The Cloud Radiative Effect - Probably should Change.. What is the Climate Sensitivity?}

Due to a combination of changing computing power and differing objectives, the sensitivity of Earth's climate to 
perturbations has been estimated in many different ways over the past few decades.  
We have computed the climate feedback parameter that appears in Table 1 with multiple methods.    
The climate sensitivity can be estimated from the change in radiative flux at the top of the 
atmosphere after imposing a prescribed change of surface temperature.  This method of using SST changes
as a proxy for climate change was pioneered by Robert Cess and his colleagues (e.g. 1989, 1990).  
Additionally, the climate
sensitivity can be estimated from a fully coupled experiment in which the $\rm{CO}_{2}$ concentration is 
instantaneously increased relative to the pre-Industrial values.  Linear regression can then be used to 
estimate the sensitivity of the model (Gregory et al., 2004, Andrews we al., 2012, IPCC).  
This is the method used to compute the climate sensitivity reported in Andrews et al., 2012 
and used by the IPCC.  The climate sensitivity
can also be computed from simply measuring the change of surface temperature after the $\rm{CO}_{2}$ concentration
has been doubled in a fully coupled climate model following an increase of 1\% per year in the 
$\rm{CO}_{2}$ concentration.   This method is used to define the Transient Climate Response (TCR).
For a particular climate model, the only way to be sure of the climate sensitivity is to run the 
coupled model after a perturbation has been applied and wait until the model has fully 
equilibrated.  Evidence indicates that this takes thousands of years of 
simulation (Stouffer, 2004; Li et al., 2012; Paynter et al., 2018; Rugenstein et al., 2019).
Because of the limiting cost of such long experiments, determining an accurate measure 
of the climate sensitivity to perturbations from a simulation that is not fully equilibrated is an 
essential component of climate science.

\begin{table}
\begin{center}
\caption{Global mean radiative feedbacks ($\rm{W\, m^{-2} K^{-1}}$).  The cloud radiative effect (CRE) is computed as
 the clear sky minus all sky TOA flux.   The \textit{amip p8K} is not a separate experiment but
shows the feedbacks computed from the \textit{amip-p4K} when the \textit{amip-m4K} is the control experiment.}
    \begin{tabular}{*{6}{c}}
    \hline
    \hline
 Feedback & Aqua p4K & Amip m4K & Amip p4K &  Amip p8K & Amip Future   \\ \hline
    Net  $\alpha$        &   -2.13      &  -1.66          &  -1.62         & -1.64           & -1.84           \\ 
    \\
    Net CRE   & -0.27       &  0.00              & 0.07         & 0.03         & -0.11            \\  
    \\
    LW CRE   & 0.38        &  0.16              & 0.13           & 0.15        & 0.10             \\  
    \\
    SW CRE  & -0.65     &  -0.16              & -0.07          & -0.12        & -0.21            \\  
    \\
    LW CLR   & -1.96       &  -2.02            & -2.02           & -2.02       & -2.03            \\  
    \\
    SW CLR  & 0.09        & 0.36              & 0.32             & 0.34        & 0.30             \\  \hline

    \end{tabular}\par
    %\bigskip 
    \label{tab:lambda}
\end{center}
\end{table}

The DECK experiments of CMIP6 can be used to provide a measure of the climate sensitivity based on 
three different fully coupled experiments.  The methods used to determine the sensitivity are distinct, and the resulting 
differences in the measures of sensitivity reveal important details of the physics that underly the response
of the climate system to a perturbation such as increasing concentrations of carbon dioxide.  The three 
experiments are the historical simulation (\textit{historical}: years 1850-2014),  the abrupt quadrupling of 
$\rm{CO}_2$ concentration (\textit{abrupt-4xCO$_2$}: 300 years for this paper), and the experiment in 
which the concentration of $\rm{CO}_2$ is increased by 1\%  per year from the pre-industrial value of 
284.262 ppm(?units?) until the concentration is doubled  (\textit{1pctCO$_2$}: total of 150 years, doubling at year 70).  
The \textit{historical} simulation is our best attempt at using a coupled GCM to reproduce the 
observed historical record.  This has frequently been used to estimate the climate sensitivity using the 
energy budget method 
(Otto et al., 2013, Armour et al., Andrews et al., \citeA{Winton_etal_2020}; Lewis and Curry?).  
The \textit{abrupt-4xCO$_2$} experiment has been utilized by the IPCC to derive measures of the 
ECS among the CMIP5 and CMIP6 models using a linear regression of the change in TOA radiative fluxes 
and surface air temperature over the first 150 years of the response to the increase in $CO_2$ 
concentration (Gregory et al., 2004; Andrews et al., 2012).  In contrast to the abrupt change of 
$\rm{CO}_2$ concentration, the \textit{1pctCO2} is an experiment of the Earth system when 
there is a  1 \% per year increase of the $\rm{CO}_2$ concentration until it is doubled relative to the 
pre-Industrial values.  The sensitivity at the point of doubling is defined 
as the Transient Climate Response (TCR).  The TCR is the relevant measure of climate sensitivity for 
near term climate change because it is expected that the atmospheric concentrations of $\rm{CO}_2$ 
will reach (insert 2xCO2 value) around the middle of this century.

The radiative feedbacks at the TOA were computed for a collection of CMIP5 models by Ringer et al., 2014.  
The results from that study provide a convenient point of reference for our findings.  
Accordingly, we use the same method to compute the radiative feedbacks for the GFDL models in Table \ref{tab:lambda}.  
That is, for the coupled abrupt $4xCO_{2}$ experiment, feedbacks are computed as  
the slopes of a linear regression calculation using annual, global-mean top-of-atmosphere flux anomalies 
(\cite{Gregory_etal_2004}; Andrews et al., 2012).  For the atmosphere only experiments 
(\textit{amip-p4K, amip-m4K, amipFuture}, and \textit{aqua-p4K}), the feedbacks are calculated as the 
differences in TOA flux anomalies between the perturbation and control experiments, normalized 
by the change in surface air temperature.  In addition, we include values that have been computed from the 
\textit{1pctCO2} experiment.   Positive values indicate a feedback that leads to warming of the climate system.

\begin{figure}
  \centering
  %\includegraphics[width=\columnwidth]{cfmipfigs/Silvers_cfmip_CRE_SinglePanel_negativesign.pdf}
  \includegraphics[width=0.75\columnwidth]{cfmipfigs/Silvers_cfmip_CRE_3Pan_vert.pdf}
  \caption{The Cloud Radiative Effect (CRE) and CRE feedback of several CMIP6 experiments.  The top panel 
  shows (black) the CRE from two versions of the CERES-EBAF observations compared to 
  simulations of CRE using AM4.0 driven by the observed SST (\textit{amip}).  Also 
  shown in the top panel is the simulated CRE from three fully coupled CM4.0 experiments (\textit{historical} (yrs. 1950-2014),
  \textit{abrupt-4$\times$CO2} (yrs. 101-150), and \textit{1pctCO2} (yrs. 101-150)).  The middle panel shows an estimate of the 
  CRE feedback of the data in the top panels, computed by normalized according to the warming of surface air temperature over 
  the relevant time period.  The bottom panel shows the CRE feedback from four 
  CFMIP experiments (\textit{amip-p4K, amip-m4K, amip-future4K, aqua-p4K}).  The sign of the \textit{amip-m4K} experiment has 
  been reversed.}
  \label{fig:CRE_feedback}
\end{figure}

% tentative hypothesis: 
% Amip Future shows a significant difference from amip p4K between about 15S and 65S.  This should lead 
% to a CRE feedback that is quite a bit larger than for amip p4k because the two experiments are have
% nearly identical CRE's at other latitudes.  
%
% meanwhile, plotting the CRE differences from amip shows that amip-p4K and amip-m4K are not as 
% similar as initially suspected from non difference plots.  When the values of amip-m4K are reversed 
% for easier comparison with the warming experiments we see that in the SH amip-m4K is actually between
% amip-future4K and amip-p4K.  In the NH amip-m4K has the strongest CRE response of all three experiments, 
% this is true north of about 40N.  Why?  
%
% I also suspect that this is largely due to the sw_cre.  
% net_cre:  0.0; 0.07; -0.11             range of 0.18. --> amip -m4K is in the middle
%  lw_cre:   0.16;  0.13;   0.10         range of 0.06
%  sw_cre: -0.16; -0.07; -0.21         range of 0.14  --> amip -m4K is in the middle

>>>>

			% Activate to display a given date or no date
%\title{Using the Walker Circulation to connect low-level cloud changes to deep convective entrainment}
\title{Clouds and Sensitivities Across a Hierarchy of GFDL CMIP6 Models}
\authors{Levi G. Silvers, David Paynter, V. Balaji, Chris Blanton, Chris Dupuis, Jeffrey Durachta, Huan Guo, 
Isaac Held, Richard Hemmler, Jasmin G. John, Pu Lin, Colleen McHugh, Serguei Nikonov, Jeffrey Ploshay, Aparna Radhakrishnan, 
Kristopher Rand, Thomas Robinson, Hans Vahlenkamp, Chandin Wilson, Bruce Wyman, Yujin Zeng, Ming Zhao}
% Isaac Held?  Michael Winton? Songmiao Fan?  M. Daniel Schwarzkopf?

@@

\begin{figure}
  \centering
% figure generated with calip25_isccp_am4_totcld.ncl, figurenumber = 1
  \includegraphics[width=0.75\columnwidth]{cfmipfigs/ISCCP_CALIPSO_clt_correctedBiasPattern_nogrid.eps}
   \caption{Comparison of total cloud fraction from the AM4 \textit{amip} experiment with two observational 
  data sets: ISCCP (left) and CALIPSO (right).  The top row shows the observational data.  The bottom row
  shows the result of subtracting the observations from the AM4.0 data.  Numbers in the title above each 
  panel give the global mean cloud fraction (top), and the bias of global mean cloud fraction (bottom)}
  \label{fig:clt_isccp_calipso}
\end{figure}

@@

Before discussing the climate sensitivities of AM4/CM4 and how they are influenced by cloud feedbacks we assess
the clouds and radiative fluxes which are simulated by AM4.0 over the recent period of 1979-2014.  This is done 
by comparing satellite based observations to the \textit{amip} experiment that is part of GFDLs 
contribution to CMIP6.  
During the development and documentation of AM4.0 \cite{Zhao_etal_2018a, Zhao_etal_2018b}, 
(Z18a and Z18b hereafter) 
and CM4 \cite{Held_etal_2019}, specific cloud properties such as the 
cloud droplet number, cloud fraction, and the partition of cloud water into liquid or ice were noted, but 
accurately modeling the radiative fluxes at the top of the atmosphere was determined to be a 
higher priority partly because of the large observational uncertainties of those cloud properties (Z18a,b).   
Based on the initial analysis for those papers, the cloud properties of AM4.0 were determined to be
at least as consistent with observations as previous generations of GFDL models.  Z18a also noted 
that AM4.0 underestimates the liquid droplet number over oceans and downwind of the aerosol emission 
sources.   
Comparisons of AM4.0 with observations from ISCCP and MODIS showed an underestimate of 
the optically thin low clouds and the mid-level clouds (Z18a, Figure 10).  
However, the observational uncertainty for some of these properties prohibited firm conclusions.  This paper
uses additional observations and a more thorough analysis of the clouds in AM4/CM4 to better determine 
the quality of the cloud simulations.        
In this section we first compare the simulated and observed cloud fraction.  We then compare  
the simulated TOA radiative fluxes from AM4.0 to those observed by CERES.    
Throughout the paper we utilize what has come to be known as the Cloud Radiative Effect (CRE), calculated 
as the difference between the full-sky and clear-sky radiative fluxes.  Earlier literature \cite<e.g.,>[]{Cess_etal_1989, Soden_etal_2004} often referred to the CRE as the Cloud 
Radiative Forcing (CRF).

The CRE is a simple and effective tool for evaluating the influence of 
clouds on the climate.  It is well known (Soden et al., 2004) that a change in the CRE is not a direct measure of the cloud 
feedback, and that non-cloud related values of CRE can occur.  However, there are many examples of 
the productive use of the CRE as a way to diagnose cloud effects (e.g. Cess et al., 1989; Webb et al., 2001; 
\citeA{Ringer_etal_2014}; Andrews et al., 2015), it is simple to calculate, and
it allows for direct comparisons with previous studies.  The pure cloud feedback usually corresponds 
to a positive shift of the change in CRE values by roughly 0.3 $\rm{W m^{-2}K^{-1}}$ (Soden et al., 2004).  The total cloud 
feedback of a model is therefore more positive than will be implied by the CRE.

\begin{figure}
  \centering
% figure generated with calip25_isccp_am4_totcld.ncl, figurenumber = 3
  \includegraphics[width=0.75\columnwidth]{/Users/silvers/Research/Silvers_cfmip_paper/cf_misr_cldfrac_below3km.eps}
  \caption{Observed and modeled clouds below 3km.  Clouds shown are divided into thick 
  (optical depth greater than and equal to 3.6) and thin (optical depth less than 3.6) 
  clouds.  The observations (bottom) are from the MISR satellite (2000-2019) the clouds from AM4.0 (top) were derived 
  with the MISR simulator in the \textit{amip} experiment. Mean cloud fraction values for MISR 
  (AM4) are 35.6\% (21.7\%) for thick clouds and 18.7\% (6.4\%) for thin clouds}
  \label{fig:cf_misr_lowcl}
\end{figure}
% MISR observations are from March 2000 through November 2019 and came from Roger Marchand

\subsection{COSP: change title}
In order to compare the observed and modeled clouds as consistently as possible across multiple platforms 
GFDL has implemented a collection of satellite simulators which are part of the 
CFMIP (Cloud Feedback Model Intercomparison Project; see \cite{Webb_etal_2017}).  Satellite simulator
are software tools that are used in coordination with GCMs to allow for model variables 
to be translated to variables that are more directly comparable with what satellites observe.  
These simulators are often referred to as the CFMIP Observation Simulator Package (COSP); for
details see Bodas-Salcedo et al. 2011.  This paper highlights simulators of the algorithms for 
four satellite datasets: ISCCP (the International Satellite Cloud Climatology Project), 
MODIS (the Moderate Resolution Imaging Spectroradiometer), 
MISR (the Multiangle Imaging Spectro-Radiometer), and 
CALIPSO (the Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation; 
Chepfer et al., 2010; Cesana and Chepfer, 2013).  By using multiple complimentary satellite 
datasets and simulators we can benefit from the different strengths of each satellite while largely 
minimizing the known problems and discrepancies of the observational data sets.

\begin{figure}
  \centering
% figure generated with cloud_table_AM4fig.ncl
  \includegraphics[width=0.98\columnwidth]{cfmipfigs/Silvers_cfmip_MISRv7_cloudtable.pdf}
  \caption{Joint histograms of cloud fraction (colors and numbers (\%)) showing the variations of cloud fraction 
  as a function of cloud top height and optical depth (tau).
  The top panels show observations from MISR, bottom panels show MISR simulator output from the 
  \textit{amip} experiment.  Shown from left to right are the mean values for the full globe, between 30S-60S,
  between 30S-30N, and between 30N-60N.}
  \label{fig:misr_sim_vs_mod}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[angle=90,width=0.85\columnwidth]{cfmipfigs/calip_9pan_amip_obs_vs_AM4.eps}
  \caption{Cloud Fraction observed by CALIPSO (2007-2016) and modeled with the CALIPSO satelite simulator 
  (amip experiment).
  Dashed line shows meridional profiles of AM4.0 cloud fraction and solid line shows
  the meridional profile of observed cloud fraction.  With only a few exceptions in the polar regions, AM4 underestimates
  clouds in all three vertical ranges with the largest underestimate being of tropical ($\pm \, 30^\circ$) low-level clouds.}
  \label{fig:calipso_9pan}
\end{figure}

@@

\section{The Role of Clouds in CM4.0, AM4.0, and Aquaplanets in Determining and Influencing Climate Sensitivity}

@@

%The range of ways that the climate sensitivity can be quantified using a single model can be revealing or confusing.  
%Recently, \cite{Zelinka_etal_2020} found the sensitivity of CM4.0 using the first 150 years of data from the \textit{abrupt-4xCO$_2$}
%experiment and the regression method of \cite{Gregory_etal_2004} to be 3.89K while \cite{Winton_etal_2020} found the sensitivity to be 5.0K using years 50-300 of the same experiment.   \citeA{Winton_etal_2020} also showed the TCR of CM4.0 to be 2.1 using the 
%\textit{1pctCO$_2$} experiment.  Additionally one could calculate the climate sensitivity by running CM4.0 for thousands of years until is equilibrated (as in \citeA{Paynter_etal_2018} for CM2 and CM3) or by using a method similar to that employed by Otto et al., 2014 to infer the sensitivity from \textit{historical} experiment.

The most recent generation of models developed at NOAA's GFDL that have participated in CMIP6 have a large range of climate sensitivities (possibly cite Held et al., Winton et al., Dunne et al.).  
One way of quantifying this sensitivity is with the climate feedback parameter, which, for the subset of GFDL models and CMIP6 experiments studied in this paper, varies between -0.46 and -2.1 $\rm{W m^{-2}K^{-1}}$ (Table 1).  
These models (AM4.0 Aquaplanet, AM4.0, CM4.0) simulate several 
distinct geophysical configurations which account for some of the range in values of the climate feedback parameter.  
However, our hypothesis is that the large range of climate sensitivities that are found within this 
group of models is in large part due to variations in the clouds (both their climatology and their feedbacks).

After briefly noting different methods of estimating the climate sensitivity we show in the first subsection how much of the difference is 
due to experimental configuration and how much is due to differences in the CRE.  
We also show (second subsection) how the cloud fraction varies among
the experiments as a function of latitude and consider the cloud distributions in the context of the observed
cloud fraction and CRE from section 3.

%\subsection{The Cloud Radiative Effect - Probably should Change.. What is the Climate Sensitivity?} 
\subsection{The Radiative Response to Warming}

%% tentative hypothesis: 
%% Amip Future shows a significant difference from amip p4K between about 15S and 65S.  This should lead 
%% to a CRE feedback that is quite a bit larger than for amip p4k because the two experiments are have
%% nearly identical CRE's at other latitudes.  
%%
%% meanwhile, plotting the CRE differences from amip shows that amip-p4K and amip-m4K are not as 
%% similar as initially suspected from non difference plots.  When the values of amip-m4K are reversed 
%% for easier comparison with the warming experiments we see that in the SH amip-m4K is actually between
%% amip-future4K and amip-p4K.  In the NH amip-m4K has the strongest CRE response of all three experiments, 
%% this is true north of about 40N.  Why?  
%%
%% I also suspect that this is largely due to the sw_cre.  
%% net_cre:  0.0; 0.07; -0.11             range of 0.18. --> amip -m4K is in the middle
%%  lw_cre:   0.16;  0.13;   0.10         range of 0.06
%%  sw_cre: -0.16; -0.07; -0.21         range of 0.14  --> amip -m4K is in the middle
%

% these numbers were computed using glbmn_toa_cfmip_fb_driver.m and glbmn_toa_cfmip_fb.m
\begin{table}
\begin{center}
\caption{Global mean radiative feedbacks ($\rm{W\, m^{-2} K^{-1}}$).  The cloud radiative effect (CRE) is computed as
 the clear sky minus all sky TOA flux.   The \textit{amip p8K} is not a separate experiment but
shows the feedbacks computed from the \textit{amip-p4K} when the \textit{amip-m4K} is the control experiment.}
    \begin{tabular}{*{6}{c}}
    \hline
    \hline
 Feedback & Aqua p4K & Amip m4K & Amip p4K &  Amip p8K & Amip Future   \\ \hline
    Net  $\alpha$        &   -2.13      &  -1.66          &  -1.62         & -1.64           & -1.84           \\ 
    \\
    Net CRE   & -0.27       &  0.00              & 0.07         & 0.03         & -0.11            \\  
    \\
    LW CRE   & 0.38        &  0.16              & 0.13           & 0.15        & 0.10             \\  
    \\
    SW CRE  & -0.65     &  -0.16              & -0.07          & -0.12        & -0.21            \\  
    \\
    LW CLR   & -1.96       &  -2.02            & -2.02           & -2.02       & -2.03            \\  
    \\
    SW CLR  & 0.09        & 0.36              & 0.32             & 0.34        & 0.30             \\  \hline

    \end{tabular}\par
    %\bigskip 
    \label{tab:lambda_AGCM}
\end{center}
\end{table}

Due to a combination of changing computing power and differing objectives, the sensitivity of Earth's climate to 
perturbations has been estimated in many different ways over the past few decades.  
We have computed the climate feedback parameter that appears in Tables 1 and 2 with multiple methods.    
The climate sensitivity can be estimated from the change in radiative flux at the top of the 
atmosphere after imposing a prescribed change of surface temperature.  This method of using SST changes
as a proxy for climate change was pioneered by Robert Cess and his colleagues (e.g. 1989, 1990) and has been widely used since.  
Additionally, the climate
sensitivity can be estimated from a fully coupled experiment in which the $\rm{CO}_{2}$ concentration is 
instantaneously increased relative to the pre-Industrial values.  Linear regression can then be used to 
estimate the sensitivity of the model (Gregory et al., 2004, Andrews we al., 2012, IPCC).  
This is the method used to compute the climate sensitivity reported in Andrews et al., 2012 
and used by the IPCC.  The Transient Climate Response (TCR), an alternate measure of climate sensitivity that is more 
relevant for the near future can be computed by measuring the change of surface temperature after the $\rm{CO}_{2}$ concentration
has been doubled in a fully coupled climate model following an increase of 1\% per year in the 
$\rm{CO}_{2}$ concentration.   
The only way to be sure of the equilibrium climate sensitivity for a particular model is to run the 
coupled model after a perturbation of $\rm{CO}_{2}$ concentration has been applied and wait until the model has fully 
equilibrated.  Evidence indicates that this takes thousands of years of 
simulation (Stouffer, 2004; Li et al., 2012; Paynter et al., 2018; Rugenstein et al., 2019).
Because of the limiting cost of such long experiments, determining an accurate measure 
of the climate sensitivity to perturbations from a simulation that is not fully equilibrated is an 
essential component of climate science.  Additional details about the efforts to determine Earth's climate sensitivity 
and relevant background can be found in Sherwood et al., 2015, Stevens et al., 2016, Knutti et al., 2017, Rugenstein et al., 2020, and Sherwood et al., 2020.

% merge the paragraph below with that above...

%The DECK experiments of CMIP6 can be used to provide estimates of the climate sensitivity based on 
%three different fully coupled experiments.  The methods used to determine the sensitivity are distinct, and the resulting 
%differences in the measures of sensitivity reveal important details of the physics that underly the response
%of the climate system to a perturbation such as increased concentrations of carbon dioxide.  The three 
%experiments are the historical simulation (\textit{historical}: years 1850-2014),  the abrupt quadrupling of 
%$\rm{CO}_2$ concentration (\textit{abrupt-4xCO$_2$}: 300 years for this paper), and the experiment in 
%which the concentration of $\rm{CO}_2$ is increased by 1\%  per year from the pre-industrial value of 
%284.262 ppm until the concentration is doubled  (\textit{1pctCO$_2$}: total of 150 years, doubling at year 70).  
%The \textit{historical} simulation is our best attempt at using a coupled GCM to reproduce the 
%observed historical record.  This has frequently been used to estimate the climate sensitivity using the 
%energy budget method 
%(Otto et al., 2013, Armour et al., Andrews et al., \citeA{Winton_etal_2020}; Lewis and Curry).  
%The \textit{abrupt-4xCO$_2$} experiment has been utilized by the IPCC to derive measures of the 
%ECS among the CMIP5 and CMIP6 models using a linear regression of the change in TOA radiative fluxes 
%and surface air temperature over the first 150 years of the response to the increase in $CO_2$ 
%concentration (Gregory et al., 2004; Andrews et al., 2012; Zelinka et al., 2020).  In contrast to the abrupt change of 
%$\rm{CO}_2$ concentration, the \textit{1pctCO2} is an experiment of the Earth system when 
%there is a  1 \% per year increase of the $\rm{CO}_2$ concentration until it is doubled (after 70 years of simulation) relative to the 
%pre-Industrial values.  The sensitivity at the point of doubling is defined 
%as the Transient Climate Response (TCR).  The TCR for CM4.0 is $2.05 (\pm0.10) K$ \cite{Winton_etal_2020} and is the relevant 
%measure of climate sensitivity for near term climate change because it is expected that the atmospheric concentrations of $\rm{CO}_2$ 
%will double relative to the pre-industrial period near the end of this century.         
%% one discussion of this is here: 
%%https://www.climate.gov/news-features/climate-qa/how-much-will-earth-warm-if-carbon-dioxide-doubles-pre-industrial-levels
%% levels rise by around 2 or 2.5 ppm per year.  They quote PI levels as 275ppm.
%
While the net feedback ($\alpha$) is a critical measure of sensitivity, the individual components of the feedback reveal 
more about the different processes that are active in the experiments.    
The radiative feedbacks at the TOA were computed for a collection of CMIP5 models by Ringer et al., 2014.  
\textit{The results from that study provide a convenient point of reference for our findings.}  
Table \ref{tab:lambda_AGCM} shows feedback values for several of the CFMIP atmosphere only experiments 
(\textit{amip-p4K, amip-m4K, amipFuture}, and \textit{aqua-p4K}).  These feedbacks are calculated as the 
differences in TOA flux anomalies between the perturbation and control experiments, normalized 
by the change in surface air temperature.   Positive values indicate a feedback that leads to warming of the climate system. 
Table \ref{tab:lambda_AGCM}  shows is that the AM4.0 aquaplanet is substantially less sensitive 
($\alpha$ is more negative) than any of the \textit{amip}-type experiments.  Approximately 0.3 $W\, m^{-2} K^{-1}$ 
of this difference is due to a weak surface-albedo feedback (SW CLR) in the aquaplanet expderiment, and the other roughly 
0.2 $W\, m^{-2} K^{-1}$ is due to a more negative net CRE.  Overall, there are strong similarities among the 
\textit{amip}-type experiments, especially in the clear sky feedback terms.  Among the \textit{amip}-type experiments, the 
patterned warming experiment (\textit{amipFuture}) is a bit (0.2 $W\, m^{-2} K^{-1}$) less sensitive, largely due to a more 
negative SW CRE response.

% Now we need to describe how the feedbacks in Table 2 have been computed...
